---
title: "Bellabeat Case Study - Marek Bem"
output: html_notebook
---

#Step 1: Install and load the packages
```{r}
install.packages('tidyverse')
install.packages('janitor')
install.packages('lubridate')
install.packages('skimr')

library(tidyverse) #wrangle data
library(janitor) #cleaning data
library(lubridate) #wrangle data attributes
library(skimr) #summary data
library(ggplot2) #visualize data
```


#Step 2: Import the databases
```{r}
daily_activity <- read.csv("dailyActivity_merged.csv")
daily_sleep <- read.csv("sleepDay_merged.csv")
weight <- read.csv("weightLogInfo_merged.csv")
hourly_step <- read.csv("hourlySteps_merged.csv")
```


#Step 3: Inspect data to see if there are any errors with formatting

```{r}
head(daily_activity)
head(daily_sleep)
head(weight)

str(daily_activity)
str(daily_sleep)
str(weight)
```


#Step 4: Check NA and duplicates / remove dupclicate

```{r}
sum(is.na(daily_activity))
sum(is.na(daily_sleep))
sum(is.na(weight))

sum(duplicated(daily_actdsleep_day))
sum(duplicated(daily_sleep))
sum(duplicated(weight))

sleep_day <- sleep_day[!duplicated(sleep_day), ]
```

#Step 5: Change format daily_activity$activity_date, daily_sleep$sleep_day and weight$date into the proper date format 

```{r}
daily_activity$activity_date <- as.Date(daily_activity$activity_date,'%m/%d/%y')
daily_sleep$sleep_day <- as.Date(daily_sleep$sleep_day, '%m/%d/%y')
weight$date <- parse_date_time(weight$date, '%m/%d/%y %H:%M:%S %p')
```

#Step 6: Removing weight$fat and daily_activity$logged_activities_distance, as it has little to no context and would not be helpful during the analysis 

```{r}
weight <- weight %>% 
  select(-c(fat))
daily_activity <-daily_activity %>% 
  select(-c(logged_activities_distance))
```


#Step 7 Add new columns which convert the current minutes of collection to hours and round it
```{r}
daily_activity$total_active_hours = round((daily_activity$very_active_minutes + daily_activity$fairly_active_minutes + daily_activity$lightly_active_minutes)/60, digits = 2)
daily_activity$sedentary_hours = round((daily_activity$sedentary_minutes)/60, digits = 2)

daily_sleep$hours_in_bed = round((daily_sleep$total_time_in_bed)/60, digits = 2)
daily_sleep$hours_asleep = round((daily_sleep$total_minutes_asleep)/60, digits = 2)
daily_sleep$time_taken_to_sleep = (daily_sleep$total_time_in_bed - daily_sleep$total_minutes_asleep)
```

#Step 8 Remove rows in which the total_active_hours & calories burned are 

```{r}
daily_activity_cleaned <- daily_activity[!(daily_activity$calories<=0),]
daily_activity_cleaned <- daily_activity_cleaned[!(daily_activity_cleaned$total_active_hours<=0.00),]
```


#Step 9: Export dateset to use in Tableau. 
```{r}
write.csv(daily_activity, file ='fitbit_daily_activity.csv')
write.csv(daily_sleep, file = 'fitbit_sleep_log.csv')
write.csv(weight, file = 'fitbit_weight.csv')
```


#Step 10 Check the average of: Steps taken, sedentary hours, very active minutes & total hours asleep.
```{r}
summary(daily_activity$total_steps)
summary(daily_activity$sedentary_hours)
summary(daily_activity$very_active_minutes)
summary(daily_sleep$hours_asleep)
```
#Percentage how many user make less then 6000 or more then 8000 steps
```
proportions(table(daily_activity$total_steps <= 6000))
proportions(table(daily_activity$total_steps >= 8000))

per_total_steps <- daily_activity %>% 
  group_by(id) %>% 
  mutate(Less6000 = if_else(total_steps <= 6000, 1, 0),
         More8000 = if_else(total_steps >= 8000, 1, 0)) %>% 
  summarize(Pct6000 = sum(Less6000) / n(),
            Pct8000 = sum(More8000) / n())
per_total_steps
```


#Step 11 Make vizualizations in R and Tableau  
```{r}
ggplot(data = daily_activity) +
  aes(x = day_of_week, y = total_distance) +
  geom_col(fill =  'orange') +
  labs(x = 'Day of week', y = 'Total distance', title = 'Total distance taken in a week')
ggsave('total_distance.png')

ggplot(data = daily_activity) +
  aes(x = day_of_week, y = very_active_minutes) +
  geom_col(fill =  'red') +
  labs(x = 'Day of week', y = 'Total very active minutes', title = 'Total activity in a week')
ggsave('total_activity.png')

ggplot(data = daily_activity) +
  aes(x = day_of_week, y = calories) +
  geom_col(fill =  'green') +
  labs(x = 'Day of week', y = 'Calories burned', title = 'Total calories burned in a week')
ggsave('total_calories.png')
```


  
#relationship between total active hours, total distance taken, and sedentary hours against calories burned 

```{r}
ggplot(data = daily_activity) +
  aes(x= total_active_hours, y = calories) +
  geom_point(color = 'red') +
  geom_smooth() +
  labs(x = 'Total active hours', y = 'Calories burned', title = 'Calories burned vs active hours')
ggsave('calories_burned_vs_active_hours.png')

ggplot(data = daily_activity) +
  aes(x= total_distance, y = calories) +
  geom_point(color = 'orange') +
  geom_smooth() +
  labs(x = 'Total steps', y = 'Calories burned', title = 'Calories burned vs total steps')
ggsave('calories_burned_vs_total_steps.png')

ggplot(data = daily_activity) +
  aes(x= sedentary_hours, y = calories) +
  geom_point(color = 'purple') +
  geom_smooth() +
  labs(x = 'Sedentary hours', y = 'Calories burned', title = 'Calories burned vs sedentary hours')
ggsave('calories_burned_vs_sedentary_hours.png')
```

#MERGE
weight_merge <- merge(daily_activity, weight, by=c('id'))

#The relationship between weight & physical activity
```{r}
ggplot(data = weight_merge) +
  aes(x = very_active_minutes, y = weight_kg) +
  geom_violin(fill = 'pink') +
  labs(x = 'Very active minutes', y = 'Weight(kg)', title = 'Relationship between weight and physical activity')
ggsave('weight_physical_activity.png')


ggplot(data = weight_merge) +
  aes(x = total_steps, y = weight_kg) +
  geom_violin(fill = 'purple') +
  labs(x = 'Total steps', y = 'Weight(kg)', title = 'Relationship between weight and physical activity')
ggsave('weight_physical_activity_steps.png')
```


